---
title: "Data Prep"
author: "Mykola Dereva"
date: "3/8/2021"
output: html_document
---

# Load libraries
```{r}
rm(list = ls())

library(tidyverse)
library(here)
library(readxl)
```


Set locale to display Ukrainian characters properly
```{r}
Sys.setlocale("LC_CTYPE", "ukrainian")
```

Load main data related to companies cost structure
Data is obtained from ukrainian statistic service (ukrstat.gov.ua)
Data is formatted in a very interesting way so I expect a lot of work to make it tidy


#  Cost Structure
```{r}
colnames <- str_c("col_", 1:28) 

data <- read_xlsx(path = here("Raw Data",
                              "Costs structure 2012-2019  +.xlsx"),
                  na = c("к/с"), 
                  skip = 16, n_max = 7744, # specific to this file
                  col_names = colnames)

```


```{r}
tail(data)
```

Data have 2 levels of aggregation I think it will be safer to parse it with
one variable in time

There is two types of missing values "к/с" and "-"
- "-" means that there is no companies 
while "к/с" that there are companies but just a few of them so 
the data is hidden out of the privacy concerns.

Therefore to differentiate those I will code 
"к/с" as NA
and "-" as 0 


UPD: it turned our that there are different types of dashes used:
```{r}
dash <- read_rds("Clean Data/dashes.rds")
dash %>% unique()
```


fill name code and year to each row:
```{r}
 data <- data %>% 
  fill(1:2, 28, .direction = "down") %>% 
  na_if("к/с") %>%  # read_xlsx did not changed it to NA
  mutate(across(where(is.character),
                ~ if_else(.x %in% dash, "0", .x) )) # change "-" to 0

```






```{r}
clean <- data %>% 
  select(28, 2:7) %>% 
  rename( "group_name" = "col_28",
          "group_code" = "col_2",
          "year"       = "col_3",
          "large"      = "col_4",
          "medium"     = "col_5",
          "small"      = "col_6",
          "micro"      = "col_7") %>% 
  pivot_longer(cols = c("large", "medium", "small", "micro"),
               names_to = "firm_size", values_to = "total_cost") %>%
  
  fill(group_name, group_code, .direction = "down") %>% 
  mutate_at(c("total_cost"), as.numeric) %>%  # convert to numeric
  mutate(year = as.integer(year))

  
head(clean)
```

```{r}
tail(clean)
```



Now we data starts to look tidy, but I have to join other variables of cost 
structure. 


Helper function to automate previous chunk

```{r}
# Here one just need to specify the columns with the data of interest
# for large, medium... company size 

extract_firms_var <- function(data,
                              group_code = "col_2",
                              year       = "col_3",
                              large, medium, small, micro,
                              var_name) {
  data %>% 
  select(group_code, year, large, medium, small, micro) %>% 
  rename( "group_code" = group_code,
          "year"       = year,
          "large"      = large,
          "medium"     = medium,
          "small"      = small,
          "micro"      = micro) %>%
  pivot_longer(cols = c("large", "medium", "small", "micro"),
               names_to = "firm_size", values_to = var_name) %>% 
  fill(group_code, .direction = "down") %>% 
  mutate_at(c(var_name), as.numeric) %>% 
  mutate(year = as.integer(year))

  
  }
                              
```


test function
```{r}
material_costs <- data %>% 
  extract_firms_var(large = "col_8", medium = "col_9",
                    small = "col_10", micro = "col_11",
                    var_name = "material_cost")

head(material_costs)
```
Seems to work fine




```{r}
depreciation <- data %>% 
  extract_firms_var(large = "col_12", medium = "col_13",
                    small = "col_14", micro = "col_15",
                    var_name = "depreciation_cost")

wages <- data %>% 
  extract_firms_var(large = "col_16", medium = "col_17",
                    small = "col_18", micro = "col_19",
                    var_name = "labour_cost")

soc_security <- data %>% 
  extract_firms_var(large = "col_20", medium = "col_21",
                    small = "col_22", micro = "col_23",
                    var_name = "social_security_cost")

other <- data %>% 
  extract_firms_var(large = "col_24", medium = "col_25",
                    small = "col_26", micro = "col_27",
                    var_name = "other_cost")
```



join to the clean dataset
```{r}
clean <- clean %>%
  left_join(material_costs) %>% 
  left_join(depreciation) %>% 
  left_join(wages) %>% 
  left_join(soc_security) %>% 
  left_join(other)

head(clean)
```


Check if the total cost equal sum of cost structure
```{r}
clean %>% 
  filter(year == 2015, firm_size == "medium") %>%
  group_by(group_code) %>% 
  mutate(sum_cost = sum(material_cost, depreciation_cost, labour_cost,
           social_security_cost, other_cost)) %>% 
  select(total_cost, sum_cost) %>% 
  head(10)
```
The numbers match. 

# Revenues

### Load data related to revenues


```{r}
colnames <- str_c("c", 1:13) 

data <- read_xlsx(path = here("Raw Data",
                              "Production Value 2012-2019  +.xlsx"),
                  na = c("к/с"),      # dont know why but this line doesnt works
                  skip = 13, n_max = 7744, # specific to this file
                  col_names = colnames)

```


```{r}
tail(data)
```



I dont know what exactly is this dash above is coded
so I assign it
```{r}
# (dash <- data$c5[7700])
```

```{r}
data <- data %>% 
  fill("c2", .direction = "down") %>% 
  na_if("к/с") %>% # read_xlsx did not changed it to NA
  mutate(across(where(is.character),
                ~ if_else(.x %in% dash, "0", .x) )) # change "-" to 0
```

```{r}
tail(data)
```


Now NAs coded properly


```{r}
revenue <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c5", medium = "c7",
                    small = "c9", micro = "c11",
                    var_name = "revenue")

head(revenue)
```


```{r}
clean <- clean %>%
  left_join(revenue)
```


# Firms number 



```{r}
colnames <- str_c("c", 1:32) 

data <- read_xlsx(path = here("Raw Data",
                              "Firms Number 2010-2019  +.xlsx"),
                  na = c("к/с"),    
                  skip = 16, n_max = 9680, # specific to this file
                  col_names = colnames)
```

```{r}
tail(data)
```

```{r}
#dash <- c(dash, data$c4[9680])
#dash
```

```{r}
data <- data %>% 
  fill("c2", .direction = "down") %>% 
  mutate(across(where(is.character),
                ~ if_else(.x %in% dash, "0", .x) )) # change "-" to 0
```


```{r}
tail(data)
```



```{r}
firms_n <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c4", medium = "c6",
                    small = "c8", micro = "c10",
                    var_name = "n_firms")
head(firms_n)
```

```{r}
clean <- clean %>%
  left_join(firms_n)
```

```{r}
head(clean)
```



# Number of Employees



```{r}
colnames <- str_c("c", 1:13) 

data <- read_xlsx(path = here("Raw Data",
                              "Number of Employers 2010-2020 +.xlsx"),
                  na = c("к/с"),      # dont know why but this line doesnt works
                  skip = 17, n_max = 10648, # specific to this file
                  col_names = colnames)

```



```{r}
tail(data)
```


```{r}
#dash <- c(dash, data$c5[10648], data$c5[10647] )
#dash <- unique(dash)
#dash
```



```{r}
data <- data %>% 
  fill("c2", .direction = "down") %>% 
  mutate(across(where(is.character),
              ~ if_else(.x %in% dash, "0", .x) )) # change "-" to 0

tail(data)
```



```{r}
employers_n <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c5", medium = "c7",
                    small = "c9", micro = "c11",
                    var_name = "employers_n")
head(employers_n)
```



```{r}
clean <- clean %>%
  left_join(employers_n)
```


# Accounting data




```{r}
colnames <- str_c("c", 1:44) 

data <- read_xlsx(path = here("Raw Data",
                              "Balance sheets 2013-2019  +.xlsx"),
                  na = c("к/с"),      # dont know why but this line doesnt works
                  skip = 15, n_max = 7745, # specific to this file
                  col_names = colnames)

tail(data)
```





```{r}
#dash <- c(dash, data$c5[7744])
#dash <- unique(dash)
#dash
```
Shit! Four different types of dashes


```{r}
data <- data %>% 
  fill("c2", .direction = "down") %>%
  na_if("к/c") %>% 
  mutate(across(where(is.character),
              ~ if_else(.x %in% dash, "0", .x) ))

tail(data)
```



```{r}
fixed_assets <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c5", medium = "c6",
                    small = "c7", micro = "c8",
                    var_name = "fixed_assets")
head(fixed_assets)
```

```{r}
current_assets <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c10", medium = "c11",
                    small = "c12", micro = "c13",
                    var_name = "current_assets")
head(current_assets)
```

```{r}
fixed_assets_and_leaving_group <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c15", medium = "c16",
                    small = "c17", micro = "c18",
                    var_name = "fixed_assets_and_leaving_group")
head(fixed_assets_and_leaving_group)
```


```{r}
equity <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c20", medium = "c21",
                    small = "c22", micro = "c23",
                    var_name = "equity")
head(equity)
```


```{r}
longterm_liabilities <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c25", medium = "c26",
                    small = "c27", micro = "c28",
                    var_name = "longterm_liabilities")
head(longterm_liabilities)
```



```{r}
current_liabilities <- data %>% 
  extract_firms_var(group_code = "c2", year = "c3",
                    large = "c30", medium = "c31",
                    small = "c32", micro = "c33",
                    var_name = "current_liabilities")
head(current_liabilities)
```

join to the clean dataset
```{r}
clean <- clean %>%
  left_join(fixed_assets) %>% 
  left_join(current_assets) %>% 
  left_join(fixed_assets_and_leaving_group) %>% 
  left_join(equity) %>% 
  left_join(longterm_liabilities) %>% 
  left_join(current_liabilities)

head(clean)
```

```{r}
#write_rds(dash,  here("Clean Data", "dashes.rds"))
```


```{r}
write_csv(x = clean,
  here("Clean Data", "full_clean_data.csv")
  )
```



I think that this is the all data I have 
Now I will have to subset only food processors

# Food Processors subset 

```{r}
food_processors <- read_excel(
  here("Raw Data", "Codes Food Producers.xlsx"),
  col_names = "c1",
  col_types = "numeric") %>%        ## there is some strange rounding 
  fill(c1, .direction = "down") %>% ## when data imported as a text
  mutate(c1 = as.character(c1)) %>% 
  pull() %>% 
  unique()

food_processors
```


```{r}
clean_food_processors <- clean %>% 
  filter(group_code %in% food_processors)
```

```{r}
write_csv(clean_food_processors,
          here("Clean Data", "clean_food_processors.csv"))
```







Check how many NAs in each group. 
To know which we can use for further analysis 

```{r}

processors_na <- clean_food_processors %>%
  group_by(group_code, group_name, firm_size) %>% 
  summarise_if(is.numeric, ~ sum(is.na(.x))) %>%
  mutate(total_na = rowSums(across(where(is.numeric))))

```

```{r}
write_csv(processors_na,
          here("Clean Data", "NA_summary.csv"))
```


